---
output: word_document
---
Github repository: <https://github.com/bredowmax/matching>

### Load & optionally install required packages
- matchingR: Runs DAA and solves the Stable Marriage Problem by matching up agents in stable pairs
- dplyr:  Allows quick filtering in large datasets
- ggplot2: Renders graphs with a professional look
- rmarkdown: Export R code to Word
- ineq: Package that allows to display inequality
```{r, message = FALSE, warning = FALSE}
requiredPackages = c("matchingR", "dplyr", "ggplot2", "rmarkdown", "ineq")
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p,character.only = TRUE)
}
```

### Assumptions to be set by researcher
Sets network size, gender distribution, preference structure, experiment iterations and random distribution variables. Preference structures are defined as follows:

- *independent/subjective*: Each preference is generated randomly and has no correlation to any other preference. It can be phrased as: "Beauty lies in the eyes of the beholder"

- *correlated/objective/horizontal*: Perfectly correlated preferences indicate that all agents agree on a score. In this model, correlated preferenced are created through the average of all individually created preferences. A real life example is a unanimously valid beauty standard

- *intercorrelated/homolog/vertical*: Certain characteristics attract agents to each other, for example a low distance, a preferable age/height/height constellation, or common interests
```{r, message = FALSE, warning = FALSE}
nmen <- 60 #number of men
nwomen <- 40 #number of women
prefIndependent <- 1 #weight of subjective/independent preferences
prefCollective <- 1 #weight of collective/correlated preferences
prefHomolog <- 1 #weight of homolog/intercorrelated preferences
repetitions <- 1 #how many times the experiment is repeated with new set of random preferences
mn <- 6.3 #average preference
sd <- 1.9 #standard deviation
```

### Automatically set parameters
Only change if you know what you are doing
```{r, message = FALSE, warning = FALSE}
seed = 1 #sets seed for reproducability
seedMin=seed
seedMax=seed+repetitions
setSize=nmen+nwomen
sumPref=prefIndependent+prefCollective+prefHomolog
wI=prefIndependent/sumPref
wC=prefCollective/sumPref
wH=prefHomolog/sumPref
RLIST=list() #Saves all matrices on a list to calculate average values at the end, like mean, sd, ...
WealthMenEachIteration=list()
WealthWomenEachIteration=list()
fo_avg_W=c()
mo_avg_W=c()
fo_avg_M=c()
mo_avg_M=c()
```

### Defining Function to generate all preferences in one iteration
Each agent has independent, correlated or intercorrelated preferences for each agent of the other gender. The mix of these has been defined by the researcher in above code chunk. This model works with cardinal preferences, which allows not only to see who is preferred, but also by how much
```{r, message = FALSE, warning = FALSE}
generatePreferences <- function(nmen, nwomen, mn, sd, wI, wC, wH) {
  #generate individual cardinal preference matrices
  iUm <- matrix(rnorm(nmen*nwomen, mn, sd), nrow=nwomen, ncol=nmen) #independent pref. by men for women
  iUw <- matrix(rnorm(nwomen*nmen, mn, sd), nrow=nmen, ncol=nwomen) #independent pref. by women for men
  
  #Calculate attractiveness score of each agent
  Aw <- rowMeans(iUm) #average rating for each woman
  Am <- rowMeans(iUw) #average rating for each man
  
  #generate homolog preferences
  hUm <- matrix(rnorm(nmen*nwomen, mn, sd), nrow=nwomen, ncol=nmen) #homolog pref. by men for women
  hUw <- t(hUm) #homolog pref. by women for men (like each other exactly the same)
  
  #Calculate preferences stated by each agent
  uM <<- (wI*iUm + wC*Aw + wH*hUm) #weighted pref. by men for women
  uW <<- (wI*iUw + wC*Am + wH*hUw) #weighted pref. by women for men
  
  #sender
  senderMen = t(t(rep(1:nmen, each = nwomen)))
  senderWomen = t(t(rep(1:nwomen, each = nmen)))
  senderList = c(senderMen, senderWomen)
  
  #receiver
  receiverWomen = rep(1:nmen, times = nwomen)
  receiverMen = rep((nmen+1):(nmen+nwomen), times = nmen)
  receiverList = c(receiverWomen, receiverMen)
  
  #preferences
  preferenceList <- c(as.list(uM), as.list(uW))
  preferences = data.frame(cbind(senderList,receiverList,preferenceList))
  names(preferences) = c("swiperID", "targetID", "utility")
  return(preferences)
}
```

### Run Experiment for x iterations
- Each experiment creates a new dating network
- Each network is two-sided and consists of n men and m women
- Each agent has random preferences for each agent of the other gender
- The preference structure is defined in the assumptions. Preferences can be independent, correlated or inter-correlated
  + The Deferred Acceptance Algorithm (DAA) solves the Stable Marriage Problem and matches up all agents in stable pairs
  + The DAA runs in two variants, which result in both the female-optimal (=male-pessimal) and the male-optimal (=female-pessimal) stable matching
  + If both results coincide, only one stable matching exists, which is optimal (and pessimal) for both genders
```{r, message = FALSE, warning = FALSE}
for (i in seedMin:seedMax){
  #set seed
  set.seed(i)
  
  #Generate Profiles
  preferences=generatePreferences(nmen, nwomen, mn, sd, wI, wC, wH)
  
  #female-optimal matching
  resultsW <<- galeShapley.marriageMarket(uW, uM) 
  galeShapley.checkStability(uW, uM, resultsW$proposals, resultsW$engagements)
  #calculate average utility of each matched agent
  resultsWW <<- diag(uW[resultsW$proposals,1:nwomen])
  uWWav <<- mean(resultsWW, na.rm=TRUE)
  resultsMW <<- diag(uM[resultsW$engagements,1:nmen])
  uMWav <<- mean(resultsMW, na.rm=TRUE)
  
  #male-optimal matching
  resultsM <- galeShapley.marriageMarket(uM, uW) 
  check <- galeShapley.checkStability(uM, uW, resultsM$proposals, resultsM$engagements)
  #calculate average utility of each matched agent
  resultsWM <<- diag(uW[resultsM$engagements,1:nwomen])
  uWMav <<- mean(resultsWM, na.rm=TRUE)
  resultsMM <<- diag(uM[resultsM$proposals,1:nmen])
  uMMav <<- mean(resultsMM, na.rm=TRUE)
  
  #Create results
  #Gini Coefficient female-optimal
  wealthW = c(pnorm(resultsWW,mean=mn,sd=sd), pnorm(resultsMW,mean=mn,sd=sd))
  wealthW[is.na(wealthW)] <- 0
  #Gini Coefficient male-optimal
  wealthM = c(pnorm(resultsMM,mean=mn,sd=sd), pnorm(resultsWM,mean=mn,sd=sd))
  wealthM[is.na(wealthM)] <- 0
  
  #Define result matrix
  #resultsUtility = matrix(c(min(resultsWW, na.rm = TRUE), uWWav, max(resultsWW, na.rm = TRUE), min(resultsWM, na.rm = TRUE), uWMav, max(resultsWM, na.rm = TRUE), min(resultsMW, na.rm = TRUE), uMWav, max(resultsMW, na.rm = TRUE), min(resultsMM, na.rm = TRUE), uMMav, max(resultsMM, na.rm = TRUE), rep(round(100*ineq(wealthW,type="Gini"),digits=0),3), rep(round(100*ineq(wealthM,type="Gini"), digits = 0),3)), nrow = 6, ncol = 3)
  resultsPercentile = matrix(c(pnorm(min(resultsWW, na.rm = TRUE),mean=mn,sd=sd), pnorm(uWWav,mean=mn,sd=sd), pnorm(max(resultsWW, na.rm = TRUE),mean=mn,sd=sd), pnorm(min(resultsWM, na.rm = TRUE),mean=mn,sd=sd), pnorm(uWMav,mean=mn,sd=sd), pnorm(max(resultsWM, na.rm = TRUE),mean=mn,sd=sd), pnorm(min(resultsMW, na.rm = TRUE),mean=mn,sd=sd), pnorm(uMWav,mean=mn,sd=sd), pnorm(max(resultsMW, na.rm = TRUE),mean=mn,sd=sd), pnorm(min(resultsMM, na.rm = TRUE),mean=mn,sd=sd), pnorm(uMMav,mean=mn,sd=sd), pnorm(max(resultsMM, na.rm = TRUE),mean=mn,sd=sd), rep(ineq(wealthW,type="Gini"),3), rep(ineq(wealthM,type="Gini"),3)), nrow = 6, ncol = 3)
  rownames(resultsPercentile) = c("fo min", "fo avg", "fo max", "mo min", "mo avg", "mo max")
  colnames(resultsPercentile) = c("Match Percentile for Women", "Match Percentile for Men", "Gini Coeff.")
  
  fo_avg_W=c(fo_avg_W,resultsPercentile[2,1])
  fo_avg_M=c(fo_avg_M,resultsPercentile[2,2])
  mo_avg_W=c(mo_avg_W,resultsPercentile[5,1])
  mo_avg_M=c(mo_avg_M,resultsPercentile[5,2])
  RLIST[[i]]=resultsPercentile
  WealthMenEachIteration[[i]]=wealthM
  WealthWomenEachIteration[[i]]=wealthW
}
```

### Results
Summarizes all x iterations and builds average of the aggregate utility over all experiment iterations. Aggregate Utility for each gender is displayed as percentile. The higher the percentile, the better. 
Example: On average, women liked their allocated match better than 75% of the other men if their percentile = 0.75
```{r, message = FALSE, warning = FALSE}
WealthMenAcrossExperiments=array(unlist(WealthMenEachIteration),c(nmen,2,repetitions+1))
WealthWomenAcrossExperiments=array(unlist(WealthWomenEachIteration),c(nwomen,2,repetitions+1))
arr=array(unlist(RLIST),c(6,3,repetitions+1))

AVERAGE=apply(arr , 1:2 , mean )
row.names(AVERAGE)=row.names(resultsPercentile)
colnames(AVERAGE)=colnames(resultsPercentile)
AVERAGE

SD=apply(arr , 1:2 , sd)
row.names(SD)=row.names(resultsPercentile)
colnames(SD)=colnames(resultsPercentile)
SD
```

### Prepare for Graphs
```{r, message = FALSE, warning = FALSE}
#TOPLOT
DF=data.frame(avg=c(fo_avg_W,fo_avg_M,mo_avg_W,mo_avg_M),
              algorithm=c(rep("female-optimal",2*(repetitions+1)),rep("male-optimal",2*(repetitions+1))),
              sex=c(rep("women",repetitions+1),
                    rep("men",repetitions+1),rep("women",repetitions+1),rep("men",repetitions+1)))

#Construct 2 violin graphs
data_summary <- function(x) {
  m <- mean(x)
  ymin <- m-sd(x)
  ymax <- m+sd(x)
  return(c(y=m,ymin=ymin,ymax=ymax))
}
```

### Density Chart
Within each iteration, was utility distributed fairly across agents in both genders? Was this result stable across iterations?
```{r, message = FALSE, warning = FALSE, dpi=300}
ggplot(data = DF, aes(x = avg)) + 
  geom_density(data = filter(DF, algorithm == "female-optimal" & sex == "men"), fill = "#2595FF", color = NA, alpha = 0.3) +
  geom_density(data = filter(DF, algorithm == "male-optimal" & sex == "men"), fill = "#2595FF", color = NA, alpha = 0.1) +
  #geom_density(data = filter(DF, algorithm == "female-optimal" & sex == "women"), fill = "#FE18AC", color = NA, alpha = 0.1) +
  #geom_density(data = filter(DF, algorithm == "male-optimal" & sex == "women"), fill = "#FE18AC", color = NA, alpha = 0.1) +
  xlab("Percentile per Agent") + ylab("Frequency") + #Add axis
  labs(title = paste("Utility Distribution per Matching")) +
  theme(plot.title = element_text(face="bold", size=16, hjust=0)) #format title
```

### Violin Chart
Across all experiment iterations, what was the aggregate utility for each gender, and how did it vary between iterations?
```{r, message = FALSE, warning = FALSE, dpi=300}
ggplot(data = DF, mapping = aes(x= sex, y = avg, fill = sex)) + 
  geom_violin(data = filter(DF, algorithm == "male-optimal"), alpha = .5) + #filter for male-optimal version of the algorithm
  geom_violin(data = filter(DF, algorithm == "female-optimal"), alpha = .5) + #filter for female-optimal version of the algorithm
  #facet_wrap(~ algorithm) + #wrap both algorithm variants in one graph
  stat_summary(fun.data=data_summary) + #add average of averages
  xlab("") + ylab("Percentile") + #Add axis
  scale_fill_manual(values=c("#2595FF", "#FE18AC")) + #coloured version
  #scale_fill_grey() + theme_classic() + #BW version
  labs(title = paste("Average Utility by Gender"), caption = paste(repetitions, " experiments"), subtitle = (paste("Set size = ", setSize," || men/women (in %) = ", nmen/setSize*100, ":", nwomen/setSize*100, "\nPreference structure individual/collective/homolog (in %) = ", round(wI*100), "/", round(wC * 100), "/", round(wH * 100), sep = ""))) + #add title
  theme(plot.title = element_text(face="bold", size=16, hjust=0)) + #format title
  ylim(low=0, high=1) + #fix y limit to 0, 1 for better comparability
  theme(legend.position="none") #remove legend
```
